---
title: "BAM_ETE_logVD_Model"
author: "ehe"
date: "2024-03-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#Load in necesary packages, notable rxode2, tidyverse. 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(devtools)
library(rxode2)
library(nlmixr2)
```

#Model code for 2-cmpt BAM integrated into Viral Dynamics w efficacy parameter 
```{r}
model_BAM <- function() {
  ini({
    # PK parameters for BAM
    ICL = 0.231  # BAM clearance, L/d
    IQ = 0.281   # BAM intercompartmental clearance
    IV1 = 2.68   # BAM volume of distribution in plasma, L
    IV2 = 2.68   # BAM volume in periphery, L
    
    # Molecular weight for BAM in g/mol
    MW_BAM = 146000
    
    # IC50 for BAM efficacy calculation in nM
    IC50_BAM = 0.17
    
    # Initial conditions for Viral Dynamics model
    tp = 7                   # End of growth phase
    tf = 30                  # Time frame of interest
    logV0 = log(3332258)     # Log initial viral load
    logVmax = log(2923894138) # Log maximum viral load
    eta.a ~ 0.1              # Random effect for growth rate
    eta.B ~ 0.1              # Random effect for decay rate
  })
  
  model({
    # Convert BAM amount (in mg) to concentration in nM
    conc_A1_nM = ((A1) / IV1) * (1/1000) * (1/ MW_BAM) * 1e9
    
    # PK equations for BAM
    KEL = ICL / IV1
    K12 = IQ / IV1
    K21 = IQ / IV2
    d/dt(A1) = -KEL * A1 - K12 * A1 + K21 * A2
    d/dt(A2) = K12 * A1 - K21 * A2
    
    # Calculate efficacy for BAM
    efficacy_BAM = conc_A1_nM / (conc_A1_nM + IC50_BAM)
    
    # Viral dynamics equations incorporating BAM efficacy
    a = (logVmax - logV0) / tp + eta.a
    B = (logVmax - logV0) / (tf - tp) + eta.B
    growthPhase = (t <= tp)
    decayPhase = (t > tp)
    
    d/dt(logV) = growthPhase * a * (1-efficacy_BAM) - decayPhase * B
    logV(0)=-1
  })
 
}

```



#Model code for 2-cmpt BAM where [BAM] in lungs is used in efficacy parameter
```{r}
model_BAM_lungs <- function() {
  ini({
    # PK parameters for BAM
    ICL = 0.231  # BAM clearance, L/d
    IQ = 0.281   # BAM intercompartmental clearance
    IV1 = 2.68   # BAM volume of distribution in plasma, L
    IV2 = 2.68   # BAM volume in periphery, L
    
    # Molecular weight for BAM in g/mol
    MW_BAM = 146000
    
    # IC50 for BAM efficacy calculation in nM
    IC50_BAM = 0.17
    
    # Initial conditions for Viral Dynamics model
    tp = 7                   # End of growth phase
    tf = 30                  # Time frame of interest
    logV0 = log(3332258)     # Log initial viral load
    logVmax = log(2923894138) # Log maximum viral load
    eta.a ~ 0.1              # Random effect for growth rate
    eta.B ~ 0.1              # Random effect for decay rate
  })
  
  model({
    #Convert BAM amount (in mg) to concentration in nM
    conc_A1_nM = ((A1) / IV1) * (1/1000) * (1/ MW_BAM) * 1e9
    #[BAM] in lungs is a function of 15% [BAM] in plasma 
    BAM_lungs = .15 * (conc_A1_nM) 
    
    # PK equations for BAM
    KEL = ICL / IV1
    K12 = IQ / IV1
    K21 = IQ / IV2
    d/dt(A1) = -KEL * A1 - K12 * A1 + K21 * A2
    d/dt(A2) = K12 * A1 - K21 * A2
    
    # Calculate efficacy for BAM
    efficacy_BAM = BAM_lungs / (BAM_lungs + IC50_BAM)
    
    # Viral dynamics equations incorporating BAM efficacy
    a = (logVmax - logV0) / tp + eta.a
    B = (logVmax - logV0) / (tf - tp) + eta.B
    growthPhase = (t <= tp)
    decayPhase = (t > tp)
    
    d/dt(logV) = growthPhase * a * (1-efficacy_BAM) - decayPhase * B
    logV(0)=-1
  })
 
}

```


#Dosing structure for BAM 
```{r}
 dose_bam_trial1<- et(amountUnits="mg", timeUnits="day") %>% 
    et(dose=700, time=0.01, cmt="A1")  %>% 
          et(c(0.01, 1/24, 1/12, 0.25, 0.5, 1, 2, 3, 4, 5, 6, 7, 8,9,10,11,12,13,14, 21, 28)) # Example of strategic time points in days
```

#Solve original model where [lung] was a function of the peripheral compartment, A2. 
```{r}
bam_vd_trial_df <- model_BAM %>% 
        rxSolve(dose_bam_trial1)
head(bam_vd_trial_df_lungs)
```

#Applied constant of 15% for the amount of BAM in the lungs as a function of [BAM] in central cmpt
```{r}
bam_vd_trial_df_lungs <- model_BAM_lungs %>% 
        rxSolve(dose_bam_trial1)
head(bam_vd_trial_df_lungs)
```

#Plots where A1 is used for efficacy parameter calculation 
```{r}
bam_vd_trial_df_lungs %>% 
  ggplot(aes(x=time, y=logV)) +
  geom_line() +
  labs(title="Viral Dynamics with BAM", x="Time (days)", y="Log Viral Load") +
        scale_y_continuous(breaks=c(0,-2,-4,-6,-8,-10,-12,-14,-16))+
  theme_minimal()

#Plot conc_A1_nM over time from bam_vd_trial_df
bam_vd_trial_df_lungs %>% 
  ggplot(aes(x=time, y=BAM_lungs)) +
  geom_line() +
  labs(title="BAM Concentration in Lungs", x="Time (days)", y="Concentration (nM)") +
        scale_y_log10()+
  theme_minimal()

```
#Plots where A2 (as a function of A1) is used for efficacy parameter calculation 
```{r}
#Plot viral load over time from bam_vd_trial_df_lungs
bam_vd_trial_df_lungs %>% 
  ggplot(aes(x=time, y=logV)) +
  geom_line() +
  labs(title="Viral Load in lungs after BAM simulated to affect viral growth phase", 
       x="Time (days)", y="Log Viral Load") +
        scale_y_continuous(breaks=c(0,-2,-4,-6,-8,-10,-12,-14,-16))+
        scale_x_continuous(breaks=c(0,3,7,10,14,21,28))+
  theme_minimal()

```
#2-cmpt model of BAM where efficacy affects viral clearance instead of replication
```{r}
model_BAM_lungs_elim <- function() {
  ini({
    # PK parameters for BAM
    ICL = 0.231  # BAM clearance, L/d
    IQ = 0.281   # BAM intercompartmental clearance
    IV1 = 2.68   # BAM volume of distribution in plasma, L
    IV2 = 2.68   # BAM volume in periphery, L
    
    # Molecular weight for BAM in g/mol
    MW_BAM = 146000
    
    # IC50 for BAM efficacy calculation in nM
    IC50_BAM = 0.17
    
    # Initial conditions for Viral Dynamics model
    tp = 7                   # End of growth phase
    tf = 30                  # Time frame of interest
    logV0 = log(3332258)     # Log initial viral load
    logVmax = log(2923894138) # Log maximum viral load
    eta.a ~ 0.1              # Random effect for growth rate
    eta.B ~ 0.1              # Random effect for decay rate
  })
  
  model({
    #Convert BAM amount (in mg) in LUNGS to concentration in nM
    conc_A1_nM = ((A1) / IV1) * (1/1000) * (1/ MW_BAM) * 1e9
    #[BAM] in lungs is a function of 15% [BAM] in plasma 
    BAM_lungs = .15 * (conc_A1_nM) 
    
    # PK equations for BAM
    KEL = ICL / IV1
    K12 = IQ / IV1
    K21 = IQ / IV2
    d/dt(A1) = -KEL * A1 - K12 * A1 + K21 * A2
    d/dt(A2) = K12 * A1 - K21 * A2
    
    # Calculate efficacy for BAM
    efficacy_BAM = BAM_lungs / (BAM_lungs + IC50_BAM)
    
    # Viral dynamics equations incorporating BAM efficacy
    a = (logVmax - logV0) / tp + eta.a
    B = (logVmax - logV0) / (tf - tp) + eta.B
    growthPhase = (t <= tp)
    decayPhase = (t > tp)
    
    d/dt(logV) = growthPhase * a - decayPhase * B*(1-efficacy_BAM)
    logV(0)=-1
  })
 
}
```


```{r}
bam_vd_trial_ELIM_df <- model_BAM_lungs_elim %>% 
        rxSolve(dose_bam_trial1)
head(bam_vd_trial_ELIM_df)
```

#Plot of BAM-2cmpt model where efficacy affects viral clearance instead of replication
```{r}
bam_vd_trial_ELIM_df %>% 
  ggplot(aes(x=time, y=logV)) +
  geom_line() +
  labs(title="Viral Load in lungs after BAM simulated to affect viral decay phase", 
       x="Time (days)", y="Log Viral Load") +
        scale_y_continuous(breaks=c(0,2,2.5,3,3.5))+
        scale_x_continuous(breaks=c(0,3,7,10,14,21,28))+
  theme_minimal()
```


#2-cmpt model for ETE where [ETE] in lungs included in efficacy parameter integrated into viral dynamics model 
```{r}
model_ETE_lungs <- function() {
  ini({
    # PK parameters for ETE
    ICL = 0.111  # ETE clearance, L/d
    IQ = 0.308   # ETE intercompartmental clearance
    IV1 = 2.45   # ETE volume of distribution in plasma, L
    IV2 = 2.18   # ETE volume in periphery, L
    
    # Molecular weight for ETE in g/mol
    MW_ETE = 145000
    
    # IC50 for ETE efficacy calculation in nM
    IC50_ETE = 0.32
    
    # Initial conditions for Viral Dynamics model
    tp = 7                   # End of growth phase
    tf = 30                  # Time frame of interest
    logV0 = log(3332258)     # Log initial viral load
    logVmax = log(2923894138) # Log maximum viral load
    eta.a ~ 0.1              # Random effect for growth rate
    eta.B ~ 0.1              # Random effect for decay rate
  })
  
  model({
    # Convert ETE amount (in mg) in LUNGS to concentration in nM
    conc_A1_nM_lungs = ((A1) / IV1) * (1/1000) * (1/ MW_ETE) * 1e9;
    ETE_lungs = .15*(conc_A1_nM_lungs);
    
    
    # PK equations for ETE
    KEL = ICL / IV1
    K12 = IQ / IV1
    K21 = IQ / IV2
    d/dt(A1) = -KEL * A1 - K12 * A1 + K21 * A2
    d/dt(A2) = K12 * A1 - K21 * A2
    
    # Calculate efficacy for ETE
    efficacy_ETE = ETE_lungs / (ETE_lungs + IC50_ETE)
    
    # Viral dynamics equations incorporating ETE efficacy
    a = (logVmax - logV0) / tp + eta.a
    B = (logVmax - logV0) / (tf - tp) + eta.B
    growthPhase = (t <= tp)
    decayPhase = (t > tp)
    
    d/dt(logV) = growthPhase * a * (1-efficacy_ETE) - decayPhase * B
    logV(0)=-1
  })
 
}

```

#Dosing structure for ETE 
```{r}
dose_ete_trial1<- et(amountUnits="mg", timeUnits="day") %>% 
    et(dose=1400, time=0.01, cmt="A1")  %>% 
          et(c(0.01, 1/24, 1/12, 0.25, 0.5, 1, 2, 3, 4, 5, 6, 7, 8,9,10,11,12,13,14, 21, 28))
```


```{r}
ete_vd_trial_df_lungs <- model_ETE_lungs %>% 
        rxSolve(dose_ete_trial1)
```
#Plot of logV over time for model_ETE_lungs
```{r}
ete_vd_trial_df_lungs %>% 
  ggplot(aes(x=time, y=logV)) +
  geom_line() +
  labs(title="Viral Load in lungs after ETE simulated to affect viral growth phase", 
       x="Time (days)", y="Log Viral Load") +
        scale_y_continuous(breaks=c(0,-2,-4,-6,-8,-10,-12,-14,-16))+
        scale_x_continuous(breaks=c(0,3,7,10,14,21,28))+
  theme_minimal()

```
#Perform Model_Type Beta for ETE to obtain graphs for slide 
```{r}
model_ETE_lungs_elim <- function() {
  ini({
    # PK parameters for ETE
    ICL = 0.111  # ETE clearance, L/d
    IQ = 0.308   # ETE intercompartmental clearance
    IV1 = 2.45   # ETE volume of distribution in plasma, L
    IV2 = 2.18   # ETE volume in periphery, L
    prop.err <- 0.1                        # Proportional error for log-transformed data
    eta.a <- 0.1            
    eta.B <- 0.1
    
    # Molecular weight for ETE in g/mol
    MW_ETE = 145000
    
    # IC50 for ETE efficacy calculation in nM
    IC50_ETE = 0.32
    
    # Initial conditions for Viral Dynamics model
    tp = 7                   # End of growth phase
    tf = 30                  # Time frame of interest
    logV0 = log(3332258)     # Log initial viral load
    logVmax = log(2923894138) # Log maximum viral load
  })
  
  model({
    # Convert ETE amount (in mg) in LUNGS to concentration in nM
    conc_A1_nM_lungs = ((A1) / IV1) * (1/1000) * (1/ MW_ETE) * 1e9;
    ETE_lungs = .15*(conc_A1_nM_lungs);
    eta.a ~ 0.1
    eta.B ~ 0.1
    
    # PK equations for ETE
    KEL = ICL / IV1
    K12 = IQ / IV1
    K21 = IQ / IV2
    d/dt(A1) = -KEL * A1 - K12 * A1 + K21 * A2
    d/dt(A2) = K12 * A1 - K21 * A2
    
    # Calculate efficacy for ETE
    efficacy_ETE = ETE_lungs / (ETE_lungs + IC50_ETE)
    
    # Viral dynamics equations incorporating ETE efficacy
    a = (logVmax - logV0) / tp + eta.a
    B = (logVmax - logV0) / (tf - tp) + eta.B
    growthPhase = (t <= tp)
    decayPhase = (t > tp)
    
    d/dt(logV) = growthPhase * a - decayPhase * B* (1+efficacy_ETE)
    logV(0)=-1
    logV ~ prop(prop.err)
  })
 
}
```

#Solve the model to obtain df
```{r}
ete_vd_trial_df_lungs_elim <- model_ETE_lungs_elim %>% 
        rxSolve(dose_ete_trial1)
```

#Plot of logV over time for model_ETE_lungs_elim
```{r}
ete_vd_trial_df_lungs_elim %>% 
  ggplot(aes(x=time, y=logV)) +
  geom_line() +
  labs(title="Viral Load in lungs after ETE simulated to affect viral decay phase", 
       x="Time (days)", y="Log Viral Load") +
         scale_y_continuous(breaks=c(0,3.6,-10,-17))+
        scale_x_continuous(breaks=c(0,3,7,10,14,21,28))+
  theme_minimal()

```

#Use nlmxr to estimate the IIV for parameters alpha and beta... or should this be applied to all parameters, including that of the 2-cmpt pk/pd immunotherapy model?
```{r}
#ete_model_beta_df_sim <- ete_vd_trial_df_lungs_elim %>% 
        #mutate(DV=logV,
               #ID = )
ete_model_beta <- nlmixr2(model_ETE_lungs_elim, data=logpatient_data_all4,est="focei")
print(ete_model_beta)
```


